Hydra Battles and Cie

### 2021-08-19 doc/part-hydras.tex
conversion to Alectryon (to be continued)
conversion to Alectryon (chapter 4, to be continued)

\begin{Coqsrc}
Lemma plus_nf : forall a,  nf a -> forall b, nf b -> nf (a + b).
Proof.
 (* ... *)

Instance plus (alpha beta : E0) : E0.
Proof.
  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
    apply plus_nf; apply cnf_ok.
Defined.

Infix "+" := plus : E0_scope.

Check omega + omega.
\end{Coqsrc}

\begin{Coqanswer}
omega + omega
     : E0
\end{Coqanswer}

@@

\begin{Coqanswer}
1 subgoal (ID 560)

  beta : T1
  n : nat
  gamma : T1
  IHbeta : nf beta -> Acc LT beta
  IHgamma : nf gamma -> Acc LT gamma
  H : nf (ocons beta n gamma)
  delta : T1
  Hdelta : delta t1< ocons beta n gamma
  ============================
  Acc LT delta
 \end{Coqanswer}

The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{ocons $\beta'$ $p'$  $\gamma'$},
where  \(\beta' \leq  \beta\) and  \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

@@

\begin{Coqsrc}
 Let Acc_strong (alpha:T1) :=
      forall n beta, 
        nf (ocons alpha n beta) -> Acc LT (ocons alpha  n beta).
\end{Coqsrc}

@@

\begin{Coqsrc}
 Lemma Acc_strong_stronger : forall alpha, 
     nf alpha -> Acc_strong  alpha -> Acc LT  alpha.
 Proof.
  intros alpha H H0; apply acc_imp with (phi0 alpha).
  - repeat split; trivial.
    + now apply lt_a_phi0_a.
  -  apply H0;  now apply single_nf.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma Acc_implies_Acc_strong : 
   forall alpha, Acc LT  alpha -> Acc_strong alpha.
\end{Coqsrc}

@@

\begin{Coqanswer}
  subgoal 1

  alpha : T1
  Aalpha : forall y : T1,  y t1< alpha -> Acc LT y
  IHalpha : forall y : T1,
       LT y alpha ->
       forall (n : nat) (beta : T1),
       nf (ocons y n beta) -> Acc LT (ocons y n beta)
  ============================
   forall (n : nat) (beta : T1),
   nf (ocons alpha n beta) -> Acc LT (ocons alpha n beta)
\end{Coqanswer}

@@

Our goal is still to prove accessibility of any well formed ordinal term.
Thanks to our previous lemmas, we are almost done.

\begin{Coqsrc}
(* A (last) structural induction *)

Theorem nf_Acc : forall alpha, nf alpha -> Acc LT  alpha.
Proof.
 induction alpha.
-  intro; apply Acc_zero.
 -  intros; eapply Acc_implies_Acc_strong;auto.
    apply IHalpha1;eauto.
    apply nf_inv1 in H; auto. 
Defined.

Corollary T1_wf : well_founded LT.
\end{Coqsrc}

@@

Definition transfinite_recursor :
 forall (P:T1 -> Type),
   (forall x:T1, 
     (forall y:T1, nf x -> nf y ->  lt y  x -> P y) -> P x) ->
    forall alpha:T1, P alpha.
Proof.
 intros; apply well_founded_induction_type with LT.
 -  exact T1_wf;auto.
 - intros. apply X. intros; apply X0. repeat split;auto. 
Defined.
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac transfinite_induction alpha :=
  pattern alpha; apply transfinite_recursor;[ | try assumption].
\end{Coqsrc}

>>>>

\input{movies/snippets/T1/plusNf}

\input{movies/snippets/E0/plusE0}
\input{movies/snippets/E0/CheckPlus}

@@

The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{ocons $\beta$ $p'$  $\gamma'$},
where    \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

@@

\input{movies/snippets/T1/AccStrongDef}

@@

\input{movies/snippets/T1/AccStrongStronger}

@@

@@

\input{movies/snippets/T1/AccImpAccStrong}

@@

Our goal is still to prove accessibility of any well formed ordinal term.
Thanks to our previous lemmas, we are almost done (by a simple structural induction!).

\input{movies/snippets/T1/nfAcc}

\input{movies/snippets/T1/T1Wf}

@@

\input{movies/snippets/T1/transfiniteRecursor}

@@

\input{movies/snippets/T1/transfiniteInduction}
